---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# laundRy

<!-- badges: start -->
[![R build status](https://github.com/UBC-MDS/laundRy/workflows/R-CMD-check/badge.svg)](https://github.com/UBC-MDS/laundRy/actions)
<!-- badges: end -->

The `laundRy` package performs many standard preprocessing techniques for Tidyverse tibbles, before use in statistical analysis and machine learning. The package functionality includes categorizing column types, handling missing data and imputation, transforming/standardizing columns and feature selection. The `laundRy` package aims to remove much of the grunt work in the typical data science workflow, allowing the analyst maximum time and energy to devote to modelling!

## Installation

You can install the released version of laundRy from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("laundRy")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("UBC-MDS/laundRy")
```

### Features
- `categorize`: This function will take in a dataframe, and output a list of lists with column types as list labels (numerical, categorical, text), and each list containing the column names associated with each column type. 

- `fill_missing`: This function takes in a dataframe and depending on user input, will either remove all rows with missing values, or will fill missing values using `mean`, `median`, or `regression` imputation. 

-  `transform_columns`: This function will take in a dataframe and apply pre-processing techniques to each column. Categorical columns will be transformed with a One Hot Encoding and numerical columns will be scaled. 

- `feature_selector`: This function takes in a dataframe which has X and y columns specified, a target task (Regression or Classification), and a maximum number of features to select. The function returns the most important features for the target task. 

## Dependencies

- TO DO

## Usage

- TODO

### LaundRy in the R ecosystem

- [mice](https://cran.r-project.org/web/packages/mice/mice.pdf) offers similar functionality for the fill_missing function, but is not integrated with a column categorizer.

- The main feature selection and preprocessing package in R is [caret](https://cran.r-project.org/web/packages/caret/caret.pdf), which carries out similar functionality to our `feature_selector` function though laundRy makes the workflow more efficient and adds imputation.

- As far as we know, there are no similar packages for Categorizing Columns and providing a list of the categorized columns. `laundRy` is the first package we are aware of to abstract away the full dataframe pre-processing workflow with a unified and simple API.

## Example

- TO DO

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(laundRy)
## basic example code
```
